\documentclass[11pt, letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=0.75in]{geometry} % Optimized margins for tables
\usepackage{booktabs}   % For professional looking tables
\usepackage{tabularx}   % For tables that automatically fit page width
\usepackage{xcolor}     % For color coding warnings
\usepackage{titlesec}   % For clean section headers
\usepackage{enumitem}   % For clean lists

% --- VISUAL SETTINGS ---
\definecolor{warningred}{RGB}{180, 0, 0}
\definecolor{forestgreen}{RGB}{0, 100, 0}

% Clean up section headers
\titleformat{\section}{\large\bfseries\uppercase}{}{0em}{}[\titlerule]
\setlength{\parindent}{0pt} % Remove paragraph indentation for a cleaner look
\setlength{\parskip}{0.5em} % Add space between paragraphs

\begin{document}

% --- TITLE ---
\begin{center}
    {\Large \textbf{OPTIMIZED SUPPLEMENT PROTOCOL}} \\
    \vspace{0.5em}
    {\small Scientific Timing \& Absorption Schedule}
\end{center}

\vspace{1em}

% --- SECTION 1: CORE RULES ---
\section{The Core Rules}

\begin{tabularx}{\textwidth}{@{} l X @{}}
    \toprule
    \textbf{Category} & \textbf{Protocol} \\
    \midrule
    \textbf{DAILY STACK} & \textbf{Vitamin D3, K2, Omega-3, Magnesium, Copper, Vitamin C.} \\
    \textit{(Do Not Cycle)} & These are biological fuels. Consistency is required for effect. \\
    \addlinespace
    \textbf{CYCLING} & \textbf{L-Theanine:} Take Mon--Fri. \textbf{Skip Sat--Sun.} \\
    \textit{(Weekdays Only)} & Prevents tolerance buildup; keeps receptors sensitive for focus. \\
    \addlinespace
    \textbf{SITUATIONAL} & \textbf{Elderberry:} Take \textbf{ONLY} when sick/flu season. \\
    \textit{(As Needed)} & Do not take daily year-round (avoids chronic immune stimulation). \\
    \addlinespace
    \textbf{ZINC NOTE} & Take \textbf{Zinc Picolinate} daily (since you take Copper). \\
    & \textit{Optional:} Skip Sundays to rest stomach lining. \\
    \bottomrule
\end{tabularx}

\vspace{0.8cm}

% --- SECTION 2: THE SCHEDULE TABLE ---
\section{Daily Schedule}

\renewcommand{\arraystretch}{1.6} % Adds breathing room to rows
\begin{tabularx}{\textwidth}{@{} p{1.8cm} p{2.5cm} >{\raggedright\arraybackslash}X >{\raggedright\arraybackslash}X @{}}
    \toprule
    \textbf{Time} & \textbf{Activity} & \textbf{Supplements To Take} & \textbf{Rationale \& Notes} \\
    \midrule
    
    % MORNING
    \textbf{07:00 AM} & \textbf{Breakfast} \newline \footnotesize{(Must have Protein)} & 
    $\bullet$ \textbf{Zinc Picolinate} \newline
    $\bullet$ \textbf{Vitamin C} \newline
    $\bullet$ \textbf{L-Theanine} (M-F) 
    & 
    \textbf{Zinc:} Taken AM to separate from Magnesium/Copper. \newline
    \textbf{Vit C:} Clears morning cortisol. \newline
    \textbf{Theanine:} Focus boost. \\
    \cmidrule(l){2-4}
    
    % NOON
    \textbf{12:00 PM} & \textbf{Lunch} & 
    \textit{None} 
    & 
    Digestive rest period. \\
    \cmidrule(l){2-4}
    
    % WORKOUT
    \textbf{05:30 PM} & \textbf{Workout} & 
    \textit{None} 
    & 
    \textcolor{warningred}{\textbf{WARNING:}} No Vit C or Elderberry here. Antioxidants blunt muscle growth signals. \\
    \cmidrule(l){2-4}
    
    % DINNER
    \textbf{07:00 PM} & \textbf{Dinner} \newline \footnotesize{(High Fat Meal)} & 
    $\bullet$ \textbf{Vitamin D3} \newline
    $\bullet$ \textbf{Vitamin K2} (MK-7) \newline
    $\bullet$ \textbf{Omega-3} (Fish Oil) \newline
    $\bullet$ \textbf{Copper} 
    & 
    \textbf{D3/K2/Omega:} Fat-soluble. Must be taken with heavy food for absorption. \newline
    \textbf{Copper:} Taken PM to avoid fighting with Zinc. \\
    \cmidrule(l){2-4}
    
    % BEDTIME
    \textbf{10:00 PM} & \textbf{Bedtime} \newline \footnotesize{(Empty Stomach)} & 
    $\bullet$ \textbf{Magnesium Glycinate} 
    & 
    Relaxes muscles and CNS. Taking it alone ensures 100\% absorption (no competition). \\
    
    \bottomrule
\end{tabularx}

\vspace{0.8cm}

% --- SECTION 3: INTERACTIONS ---
\section{Interaction Cheat Sheet}

\begin{tabularx}{\textwidth}{@{} l l X @{}}
    \toprule
    \textbf{Interaction} & \textbf{Status} & \textbf{The Rule} \\
    \midrule
    Zinc vs. Copper & \textcolor{warningred}{\textbf{ENEMIES}} & They block each other. \textbf{Split them} (Zinc AM / Copper PM). \\
    Zinc vs. Magnesium & \textcolor{warningred}{\textbf{ENEMIES}} & They compete for absorption. \textbf{Split them} (Zinc AM / Mag PM). \\
    D3 + K2 + Omega-3 & \textcolor{forestgreen}{\textbf{ALLIES}} & They require fat. \textbf{Stack them} at dinner. \\
    Vit C + Workout & \textcolor{warningred}{\textbf{BAD TIMING}} & Kills the pump. Never take antioxidants post-workout. \\
    \bottomrule
\end{tabularx}

\end{document}